---
title : "apathe: American Psychological Association Thesis<br />with R Markdown"
output: github_document
---

```{r include = FALSE}
if(!requireNamespace("depgraph", quietly = TRUE)) remotes::install_github("crsh/depgraph")
```


<!-- README.md is generated from README.Rmd. Please edit that file -->

[![CRAN/METACRAN](https://img.shields.io/cran/v/apathe?label=CRAN&logo=r)]( https://cran.r-project.org/package=apathe) [![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip) ![GitHub last commit (main)](https://img.shields.io/github/last-commit/crsh/apathe/main?label=Last%20commit&logo=github) [![R-CMD-check](https://github.com/crsh/apathe/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/crsh/apathe/actions/workflows/R-CMD-check.yaml) [![codecov](https://codecov.io/gh/crsh/apathe/branch/master/graph/badge.svg)](https://app.codecov.io/gh/crsh/apathe) [![GitHub bug issues](https://img.shields.io/github/issues/crsh/apathe/bug?label=Bugs&logo=github)](https://github.com/crsh/apathe/issues?q=is%3Aopen+is%3Aissue+label%3Abug)

**apathe** is an R package that facilitates writing computationally reproducible student theses that conform to the American Psychological Association (APA) manuscript guidelines (6th Edition).
The package provides an [R Markdown](https://rmarkdown.rstudio.com/) template that can be used with (or without) [RStudio](https://posit.co/) to create PDF documents (using the [apa6](http://www.ctan.org/pkg/apa6) LaTeX class).
**apathe** is built on top of the package [**papaja**](https://github.com/crsh/papaja) but is tailored to the specific requirements of student theses and omits some features for simplicity.

If you believe you have found a bug or would like to request a new feature, [open an issue](https://github.com/crsh/apathe/issues) on Github and provide a [minimal complete verifiable example](https://stackoverflow.com/help/minimal-reproducible-example).

## Installation

To use **apathe** you need either a recent version of [RStudio](https://posit.co/) or [pandoc](https://pandoc.org/).
You additionally need a [TeX](https://en.wikipedia.org/wiki/TeX) distribution; we recommend you use [TinyTex](https://yihui.org/tinytex/), which can be installed from within R:

```{r eval = FALSE}
if(!requireNamespace("tinytex", quietly = TRUE)) install.packages("tinytex")

tinytex::install_tinytex()
```

Using other TeX distributions is possible.
Please refer to the [**papaja** manual](https://frederikaust.com/papaja_man/introduction.html#getting-started) for detailed installation instructions.

**apathe** is not yet available on CRAN but you can install it from this GitHub repository:

```{r install-apathe-github, eval = FALSE}
# Install remotes package if necessary
if(!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")

# Install the stable development version from GitHub
remotes::install_github("crsh/apathe")
```

## Usage

Once **apathe** is installed, you can select the APA-style thesis template when creating a new R Markdown file through the RStudio menus.

To add citations, specify your bibliography-file in the YAML front matter of the document (`bibliography: my.bib`) and start citing (for details, see pandoc manual on the [citeproc extension](https://pandoc.org/MANUAL.html#extension-citations)). **apathe** supports DOI-based citations via the [`doi2cite`](https://github.com/korintje/pandoc-doi2cite/blob/main/doi2cite.lua) filter, see [**rmdfiltr** vignette](https://CRAN.R-project.org/package=rmdfiltr/vignettes/doi2cite.html). Any of the follow DOI tags can be used in the text:

- `@https://doi.org/`
- `@doi.org/`
- `@DOI:`
- `@doi:`

For example, `@doi:10.1080/01621459.2025.2516210`. When the document is rendered, `doi2cite` queries CrossRef for the bibliographic information, writes it to a local BibTeX-file and replaces the citation key by the proper BibTeX key.

This requires that the designated file `__from_DOI.bib` (it currently has to be this file name!) is added to the `bibliography` field of the YAML front matter

## Citation

Please cite **apathe** if you use it.

    `r attr(unclass(citation("apathe"))[[1]], "textVersion")`

For convenience, you can [use `cite_r()`](https://frederikaust.com/papaja_man/writing.html#citing-r-and-its-packages) or copy the reference information returned by `citation('apathe')` to your BibTeX file:

~~~bibtex

```{r echo = FALSE, results = "asis"}
print(citation("apathe"), style = "Bibtex")
```

~~~


## Contribute

[![GitHub help wanted issues](https://img.shields.io/github/issues/crsh/apathe/help%20wanted?logo=github&logoColor=%2523FFF)](https://github.com/crsh/apathe/issues?q=is%3Aopen+is%3Aissue+label%3A%22help+wanted%22) [![GitHub documentation issues](https://img.shields.io/github/issues/crsh/apathe/documentation?logo=github&logoColor=%2523FFF)](https://github.com/crsh/apathe/issues?q=is%3Aopen+is%3Aissue+label%3Adocumentation)

Like **apathe** and want to contribute?
We highly appreciate any contributions to the R package or its documentation.
Take a look at the [open issues](https://github.com/crsh/apathe/issues) if you need inspiration.

### Code of Conduct

Please note that the **apathe** project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## Package dependencies

```{r dep-plot, echo = FALSE, fig.width = 10, fig.height = 9, message = FALSE, warning = FALSE}
depgraph::plot_dependency_graph(
  pkg = "."
  , repos = "https://cran.rstudio.com/"
)
```
